\bf An Examination of Sport Climbing Competition Format and Scoring System

### 2021-08-04 manuscript/Rmd/draft.Rmd
update draft

<!-- In what percentage of the (6!)^3 orderings would the winner change if one of the 6 competitors was dropped and there were only 5 instead of 6? -->

>>>>

### Probability of Reordering in Higher Ranks
We first look at the rate at which the ranks above that of which is dropped are reordered.
```{r, echo = FALSE}
load("draft_files/reorder_all_above_probs.rdata")

sum_dat1 <- reorder_all_above %>%
  group_by(dropped_rank) %>%
  rename("Dropped Rank" = dropped_rank) %>%
  summarize("Simulations" = n()
            , "Avg. Probability" = mean(displ_prob)
            , "Median Probability" = median(displ_prob)
            , "St. Deviation" = sd(displ_prob)
            )

kable(sum_dat1)
```

### Probability of Change in Finalists

A potential event of interest is the replacement of one finalist with another, taking the chance away from the original finalist to win a medal in the finals round. Here we compute the probability that a finalist will be replaced with a non-finalist after removing another non-finalist from the rankings. Below is a table summarizing the results.

```{r, echo=FALSE}
load("draft_files/drop_qualifier_sims.rdata")

reorder_dat2 <- lapply(9:20
                  , function(x) {
                    data.frame(dropped_rank = x
                               , displ_prob = pct_displace_finalists[[x-8]]
                               )
                    }
                  ) %>%
  do.call(rbind, .)

sum_dat2 <- reorder_dat2 %>%
  group_by(dropped_rank) %>%
  rename("Dropped Rank" = dropped_rank) %>%
  summarize("Simulations" = n()
            , "Avg. Probability" = mean(displ_prob)
            , "Median Probability" = median(displ_prob)
            , "St. Deviation" = sd(displ_prob)
            )

kable(sum_dat2)

```

as can be seen, the probability is relatively low when the 9th place competitor is removed. My hypothesis is that there is enough separation between the original finalists and the 10th+ ranking competitors in the individual events that there isn't a high chance that the total rankings of the finalists can be overtaken by 10+th place competitors. Note: it might be worth computing the probability of getting a low ranking in any of the individual events, given that a competitor is a finalist (ie: $P(r_1 > 8 \text{ or } r_2 > 8 \text{ or } r_3 > 8 \text{ }|\text{ } R_T \leq 8)$).

### Probability that Medalists are Reordered or Changed

The second scenario of interest is if a medalist is displaced by another medalist, or even a non-medalist. Below are the results summarizing the probability of this scenario occurring under uniform random ranking conditions.

```{r, echo = FALSE}
# probability that medalists are displaced (changed or reordered)
load("draft_files/top3_displace_probs.rdata")

displace_probs %>%
  group_by(dropped_rank) %>%
  rename("Dropped Rank" = dropped_rank) %>%
  summarize("Simulations" = n()
            , "Avg. Probability" = mean(displ_prob)
            , "Median Probability" = median(displ_prob)
            , "St. Deviation" = sd(displ_prob)
            ) %>%
 kable()

```

### 2021-08-04 manuscript/Rmd/draft.tex
update draft

\caption{\label{tab:unnamed-chunk-4}Descriptive statistics of simulated scores for qualification and final rounds}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
round & min & Q1 & median & Q3 & max & mean & sd & n\\
\midrule
Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.84378 & 1273.47884 & 200000\\
Final & 1 & 24 & 60 & 126 & 512 & 91.19935 & 91.11065 & 80000\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-6}The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.}
\centering
\begin{tabular}[t]{rr}
\toprule
Rank & Average score\\
\midrule
1 & 36.0187\\
2 & 73.6111\\
3 & 115.3954\\
4 & 162.2263\\
5 & 216.0041\\
\addlinespace
6 & 278.1649\\
7 & 350.3272\\
8 & 434.5932\\
9 & 532.1383\\
10 & 642.3298\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-7}The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.}
\centering
\begin{tabular}[t]{rr}
\toprule
Rank & Average score\\
\midrule
1 & 9.6687\\
2 & 20.1844\\
3 & 33.3658\\
4 & 50.5734\\
5 & 75.2499\\
\addlinespace
6 & 110.7070\\
7 & 164.8258\\
8 & 265.0198\\
\bottomrule
\end{tabular}
\end{table}

@@

@@

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-9-1}

@@

\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-9}
\end{figure}

@@

@@

\caption{This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out one climber. For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results. Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-12}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-14}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5731 & 0.6760 & 0.6126 & 1.0000 & 0.5474 & 0.6878 & 0.7053\\
Speed & 0.5731 & 1.0000 & 0.3022 & 0.2174 & 0.5474 & 1.0000 & 0.2540 & 0.2947\\
Bouldering & 0.6760 & 0.3022 & 1.0000 & 0.6203 & 0.6878 & 0.2540 & 1.0000 & 0.6878\\
Lead & 0.6126 & 0.2174 & 0.6203 & 1.0000 & 0.7053 & 0.2947 & 0.6878 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-15}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.2632 & 0.3684 & 0.3474 & 1.0000 & 0.2316 & 0.5013 & 0.2526\\
Speed & 0.2632 & 1.0000 & -0.0947 & -0.3263 & 0.2316 & 1.0000 & 0.0686 & -0.4526\\
Bouldering & 0.3684 & -0.0947 & 1.0000 & 0.2000 & 0.5013 & 0.0686 & 1.0000 & 0.2902\\
Lead & 0.3474 & -0.3263 & 0.2000 & 1.0000 & 0.2526 & -0.4526 & 0.2902 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-16}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.0947 & 0.3810 & 0.2632 & 1.0000 & 0.1847 & 0.5358 & 0.5368\\
Speed & 0.0947 & 1.0000 & -0.1587 & -0.4316 & 0.1847 & 1.0000 & -0.1915 & -0.0897\\
Bouldering & 0.3810 & -0.1587 & 1.0000 & 0.2751 & 0.5358 & -0.1915 & 1.0000 & 0.5040\\
Lead & 0.2632 & -0.4316 & 0.2751 & 1.0000 & 0.5368 & -0.0897 & 0.5040 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-17}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5906 & 0.5556 & 0.6491 & 1.0000 & 0.5789 & 0.6140 & 0.5906\\
Speed & 0.5906 & 1.0000 & 0.2164 & 0.4035 & 0.5789 & 1.0000 & 0.2632 & 0.3099\\
Bouldering & 0.5556 & 0.2164 & 1.0000 & 0.2749 & 0.6140 & 0.2632 & 1.0000 & 0.3918\\
Lead & 0.6491 & 0.4035 & 0.2749 & 1.0000 & 0.5906 & 0.3099 & 0.3918 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

>>>>

\caption{\label{tab:unnamed-chunk-4}Descriptive statistics of simulated scores for qualification and final rounds}
\centering
\begin{tabular}[t]{lrrrrrrrr}
\toprule
round & min & Q1 & median & Q3 & max & mean & sd & n\\
\midrule
Qualification & 1 & 240 & 684 & 1638 & 8000 & 1158.23077 & 1269.92888 & 200000\\
Final & 1 & 24 & 60 & 126 & 512 & 91.02296 & 90.68192 & 80000\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-6}The average scores for the top 10 qualification ranks according to our simulations. A climber will secure a finalist spot if they finish in the top 8.}
\centering
\begin{tabular}[t]{rr}
\toprule
Rank & Average score\\
\midrule
1 & 35.8986\\
2 & 73.6254\\
3 & 115.0981\\
4 & 162.6459\\
5 & 216.5088\\
\addlinespace
6 & 278.7528\\
7 & 351.4263\\
8 & 435.0092\\
9 & 532.5390\\
10 & 642.8740\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-7}The average scores for all final ranks according to our simulations. Ranks 1, 2, and 3 are table positions that guarantee medalist status for climbers.}
\centering
\begin{tabular}[t]{rr}
\toprule
Rank & Average score\\
\midrule
1 & 9.6823\\
2 & 20.3044\\
3 & 33.3976\\
4 & 50.5062\\
5 & 74.9904\\
\addlinespace
6 & 110.3614\\
7 & 164.9222\\
8 & 264.0192\\
\bottomrule
\end{tabular}
\end{table}

@@

\hypertarget{probability-of-reordering-in-higher-ranks}{%
\subsubsection{Probability of Reordering in Higher
Ranks}\label{probability-of-reordering-in-higher-ranks}}

We first look at the rate at which the ranks above that of which is
dropped are reordered.

\begin{tabular}{r|r|r|r|r}
\hline
Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
\hline
4 & 100 & 0.25869 & 0.2590 & 0.0145433\\
\hline
5 & 100 & 0.34032 & 0.3405 & 0.0151263\\
\hline
6 & 100 & 0.38012 & 0.3820 & 0.0152430\\
\hline
7 & 100 & 0.40636 & 0.4060 & 0.0149582\\
\hline
8 & 100 & 0.39439 & 0.3930 & 0.0151023\\
\hline
\end{tabular}

\hypertarget{probability-of-change-in-finalists}{%
\subsubsection{Probability of Change in
Finalists}\label{probability-of-change-in-finalists}}

A potential event of interest is the replacement of one finalist with
another, taking the chance away from the original finalist to win a
medal in the finals round. Here we compute the probability that a
finalist will be replaced with a non-finalist after removing another
non-finalist from the rankings. Below is a table summarizing the
results.

\begin{tabular}{r|r|r|r|r}
\hline
Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
\hline
9 & 100 & 0.0505 & 0.050 & 0.0217597\\
\hline
10 & 100 & 0.1538 & 0.150 & 0.0318386\\
\hline
11 & 100 & 0.1603 & 0.160 & 0.0343616\\
\hline
12 & 100 & 0.1416 & 0.140 & 0.0353259\\
\hline
13 & 100 & 0.1365 & 0.140 & 0.0336162\\
\hline
14 & 100 & 0.1217 & 0.120 & 0.0344965\\
\hline
15 & 100 & 0.1213 & 0.120 & 0.0308353\\
\hline
16 & 100 & 0.1159 & 0.120 & 0.0306197\\
\hline
17 & 100 & 0.1086 & 0.105 & 0.0314312\\
\hline
18 & 100 & 0.1026 & 0.100 & 0.0329867\\
\hline
19 & 100 & 0.0886 & 0.090 & 0.0291641\\
\hline
20 & 100 & 0.0700 & 0.070 & 0.0258590\\
\hline
\end{tabular}

as can be seen, the probability is relatively low when the 9th place
competitor is removed. My hypothesis is that there is enough separation
between the original finalists and the 10th+ ranking competitors in the
individual events that there isn't a high chance that the total rankings
of the finalists can be overtaken by 10+th place competitors. Note: it
might be worth computing the probability of getting a low ranking in any
of the individual events, given that a competitor is a finalist (ie:
\(P(r_1 > 8 \text{ or } r_2 > 8 \text{ or } r_3 > 8 \text{ }|\text{ } R_T \leq 8)\)).

\hypertarget{probability-that-medalists-are-reordered-or-changed}{%
\subsubsection{Probability that Medalists are Reordered or
Changed}\label{probability-that-medalists-are-reordered-or-changed}}

The second scenario of interest is if a medalist is displaced by another
medalist, or even a non-medalist. Below are the results summarizing the
probability of this scenario occurring under uniform random ranking
conditions.

\begin{tabular}{r|r|r|r|r}
\hline
Dropped Rank & Simulations & Avg. Probability & Median Probability & St. Deviation\\
\hline
4 & 100 & 0.25872 & 0.2580 & 0.0137106\\
\hline
5 & 100 & 0.31988 & 0.3190 & 0.0161416\\
\hline
6 & 100 & 0.28741 & 0.2870 & 0.0127969\\
\hline
7 & 100 & 0.25631 & 0.2545 & 0.0145301\\
\hline
8 & 100 & 0.20631 & 0.2050 & 0.0120283\\
\hline
\end{tabular}

@@

@@

\caption{Kendall's rank correlations - 2018 World Championship, Men's and Women's Qualifications}\label{fig:unnamed-chunk-12}
\end{figure}

@@

{\centering \includegraphics{draft_files/figure-latex/unnamed-chunk-15-1}

@@

\caption{This figure illustrates the changes to the final rankings of the 2018 Youth Olympics Men's and Women's Finals when we leave out one climber. For each gender competition, each panel represents the rank of the drop-out athlete, with 0 being the original final results. Each case with a change in rank orderings is highlighted by a black panel border, and any player with a rank change is represented by a red-filled bar}\label{fig:unnamed-chunk-15}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-17}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2018 Asian Games}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5731 & 0.6760 & 0.6126 & 1.0000 & 0.5474 & 0.6878 & 0.7053\\
Speed & 0.5731 & 1.0000 & 0.3022 & 0.2174 & 0.5474 & 1.0000 & 0.2540 & 0.2947\\
Bouldering & 0.6760 & 0.3022 & 1.0000 & 0.6203 & 0.6878 & 0.2540 & 1.0000 & 0.6878\\
Lead & 0.6126 & 0.2174 & 0.6203 & 1.0000 & 0.7053 & 0.2947 & 0.6878 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-18}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2019 World Championships}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.2632 & 0.3684 & 0.3474 & 1.0000 & 0.2316 & 0.5013 & 0.2526\\
Speed & 0.2632 & 1.0000 & -0.0947 & -0.3263 & 0.2316 & 1.0000 & 0.0686 & -0.4526\\
Bouldering & 0.3684 & -0.0947 & 1.0000 & 0.2000 & 0.5013 & 0.0686 & 1.0000 & 0.2902\\
Lead & 0.3474 & -0.3263 & 0.2000 & 1.0000 & 0.2526 & -0.4526 & 0.2902 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-19}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 European Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.0947 & 0.3810 & 0.2632 & 1.0000 & 0.1847 & 0.5358 & 0.5368\\
Speed & 0.0947 & 1.0000 & -0.1587 & -0.4316 & 0.1847 & 1.0000 & -0.1915 & -0.0897\\
Bouldering & 0.3810 & -0.1587 & 1.0000 & 0.2751 & 0.5358 & -0.1915 & 1.0000 & 0.5040\\
Lead & 0.2632 & -0.4316 & 0.2751 & 1.0000 & 0.5368 & -0.0897 & 0.5040 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-20}This table shows Kendall's rank correlation coefficients between the overall and individual discipline ranks for men's and women's sport climbing qualifications at the 2020 Pan American Championship}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{4}{c}{Men} & \multicolumn{4}{c}{Women} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
  & Overall & Speed & Bouldering & Lead & Overall & Speed & Bouldering & Lead\\
\midrule
Overall & 1.0000 & 0.5906 & 0.5556 & 0.6491 & 1.0000 & 0.5789 & 0.6140 & 0.5906\\
Speed & 0.5906 & 1.0000 & 0.2164 & 0.4035 & 0.5789 & 1.0000 & 0.2632 & 0.3099\\
Bouldering & 0.5556 & 0.2164 & 1.0000 & 0.2749 & 0.6140 & 0.2632 & 1.0000 & 0.3918\\
Lead & 0.6491 & 0.4035 & 0.2749 & 1.0000 & 0.5906 & 0.3099 & 0.3918 & 1.0000\\
\bottomrule
\end{tabular}}
\end{table}
